# Using Packages {#packages}

You can think of packages as collections of R codes (and more!) that other people have created to make R more powerful or easier to use. There are many different packages that serve different functions. Regardless which package you use, you need to both install and load the package. 

In this section, we will demonstrate how to install, load, and use the `sylcount` package, which computes Flesch Reading Ease scores to estimate the readability of the items in a scale.

## Installing and Loading sylcount

To install `sylcount`, use the function `install.packages()`. Make sure you are connected to the internet. The packages will be downloaded from an R server and automatically installed. It's very common for R packages to rely on other packages to work correctly. To make sure all necessary packages are downloaded, the command below includes the specification `dependencies = T`.

```{r eval = FALSE}

install.packages("sylcount", dependencies = TRUE)

```

Even after we have installed `sylcount`, we cannot use it unless we load it. To load it, we need the `library()` function. 

```{r echo = TRUE, message = FALSE, warning = FALSE, error = FALSE}

library(sylcount)

```

Note that we only need to install each package once. But we need to load the package each time we re-start RStudio. (Think of the package like a genie that goes to sleep each time you close RStudio. If you need the genie to do stuff for you the next time around, you need to awaken it again!)

## Using sylcount

Now, suppose we have two items in our scale: 

- Assess your level of subjective well-being over the past fortnight.
- How happy have you been over the past two weeks?

### Readability of Each Item

Let's say we want to know how readable each item is. To do that, we will first store our items as individual R objects. 

```{r}

# Store the first item as item1
item1 <- " Assess your level of subjective well-being over the past fortnight."

# Store the second item as item2
item2 <- " How happy have you been over the past two weeks?"

```

Notice that we added a space after the first quotation mark for each object. Later, we will combine all the items into a single text sample. Adding a space after the first quotation mark ensures that the number of words is computed correctly when we combine the items.

Then, we will use the `readability()` function from the `sylcount` package to compute the readability of each item. 

```{r}

# Readability for item1 
readability(item1)

# Readability for item2
readability(item2)

```

The Flesch Reading Ease score is presented under the 're' column. 

Remember that the Flesch Reading Ease score ranges from 0 (difficult to understand) to 100 (easy to understand). We want to aim for a score over 60. 

### Readability of All Items

What if we want to compute the readability of all items? To do this, we first paste all the items into a single object using the `paste0()` function. 

```{r}

all_items <- paste0(item1, item2)

```

Then we run the readability check. 

```{r}

readability(all_items)

```

And that's it! Congratulations! You've learnt how to install, load, and use a package to find out the readability of items in a scale!

In the [following section](#reliability), we will use more packages (yay!) to find the reliability of a scale. 
